Tables

### 2021-08-04 analysis/article.Rmd
feat: add table 1 and table observed vs expected mortality

```{r article_setup, include=FALSE}
library("targets")

tar_load(raw_data)
tar_load(bmrk_results)
```

@@

In a retrospective cohort study we followed `r n_total` consecutive
patients, who were recruited during the evaluation process for
liver transplantation at the University Hospital of Leipzig from
`r paste(format(attr(raw_data, "period"), "%B %Y"), collapse = " to ")`.
For each patient we recorded `r ncol(raw_data)` variables.
Among them are age, sex, etiology of liver disease, complications as listed in
Table \@ref(tab:tableone) and `r length(cn_labs)` laboratory measurements.

@@

>>>>

```{r article_setup, include=FALSE}
library("targets")
library("ameld")

tar_load(bmrk_results)

tar_load(raw_data)
tar_load(labelled_meld_data)

tar_load(tbl1)
tar_load(tbl_observed_vs_expected_mortality)
```

@@

In a retrospective cohort study we followed `r n_total` consecutive
patients, who were recruited during the evaluation process for
liver transplantation at the University Hospital of Leipzig from
`r paste(format(attr(raw_data, "period"), "%B %Y"), collapse = " to ")`.
For each patient we recorded `r ncol(raw_data)` variables.
Among them are age, sex, etiology of liver disease, complications as listed in
Table \@ref(tab:tbl1) and `r length(cn_labs)` laboratory measurements.

@@

## Baseline characteristics

```{r tbl1, results = "asis", echo = FALSE, message = FALSE}
tbl1
```

<!-- TODO:
use `pattern` argument to provide "{n} ({p}%)" for categorial/dichotomos
variables after this PR is merged
https://github.com/ddsjoberg/gtsummary/pull/951
-->

```{r tbl1_helper_func, echo = FALSE}
.tbl1 <- function(variable, column = "Overall", ...) {
    id <- gtsummary::inline_text(
        tbl1, variable = all_of(variable), column = "tbl_id1"
    )
    if (column == "Overall")
        column <- "stat_0"
    gtsummary::inline_text(
        tbl1$tbls[[id]],
        variable = all_of(variable), column = all_of(column), ...
    )
}
```

In this study we analysed a cohort of `r n_analysed` patients evaluated for
liver transplantation (Tab. \@ref(tab:tbl1)).
`r sprintf("%i (%s)", sum(raw_data$Sex == "male"), gtsummary::style_percent(mean(raw_data$Sex == "male"), symbol = TRUE))`
of them were men.
The median (IQR) follow-up times was `r .tbl1("DaysAtRisk")` days.
Within the follow-up time `r .tbl1("LTx")` patients received a liver transplant
and were censored for the analysis beyond this time point.
The median (IQR) age of the cohort was `r .tbl1("Age")` years.

Most of the patients presented with cirrhosis
(`r .tbl1("Cirrhosis")`).
The remaining patients had an acute liver failure
(`r .tbl1("ALF")`) or
a chronic liver insufficiency
(`r .tbl1("CLI")`).
At evaluation time the median (IQR) MELD and MELD-Na score were
`r .tbl1("ScoreMeldUnos")` and
`r .tbl1("ScoreMeldNaUnos")`, respectively.
Within 90 days `r .tbl1("M90")` patients died.
In nearly two third of all patients alcohol was the main cause
(`r .tbl1("Ethyltoxic")`) of their end-stage liver disease.
The remaining patients had viral infections
(hepatitis B and C viruse `r .tbl1("HBV")` and `r .tbl1("HCV")`, respectively),
autoimmune processes (autoimmune hepatitis `r .tbl1("AIH")`,
primary sclerosing cirrhosis `r .tbl1("PSC")`),
primary biliary cirrhosis `r .tbl1("PBC")` or unknown causes
(non-alcoholic steatohepatitis `r .tbl1("NASH")` and
cryptogenic hepatitis `r .tbl1("Cryptogenic")`).
The most common complications among all patients were gastrointestinal
bleedings in `r .tbl1("GIB")` patients, followed by
hepatocellular carcinoma (HCC) in `r .tbl1("HCC")` and
spontaneous bacterial peritonitis (SBP) in `r .tbl1("SBP")` patients.
Just `r .tbl1("Dialysis")` patients required a renal dialysis.

According to @wiesner2003 we divided our cohort into
`r nlevels(labelled_meld_data$MeldCategory)` MELD-score risk
categories, with `r enum(table(labelled_meld_data$MeldCategory))` patients,
respectively.
As depicted in Figure \@ref(fig:observed-vs-meld-mortality-plot) and Table
\@ref(tab:tbl-observed-vs-expected-mortality) our observed 90 day mortality was
`r paste(round(range(tbl_observed_vs_expected_mortality$SMR[-1]), 1), collapse = " to ")`
times higher than the predicted one. Just in the group with the lowest MELD
scores the mortality was much lower.

```{r tbl-observed-vs-expected-mortality, results = "asis", echo = FALSE, message = FALSE}
knitr::kable(
    cbind(
        row.names(tbl_observed_vs_expected_mortality),
        tbl_observed_vs_expected_mortality
    ),
    row.names = FALSE,
    col.names =
        c(
            "MELD category",
            "Observed deaths (n)",
            "Expected deaths (n)",
            "Standardized mortality ratio (SMR)",
            "Observed mortality (%)",
            "Expected mortality (%)"
        ),
    caption = paste0(
        "Observed vs MELD-expected 90 day mortality. ",
        "MELD mortality values are taken from @wiesner2003. ",
        "All patients censored before day 90 are ignored for ",
        " the calculation of the MELD-expected deaths. ",
        "SMR, Standardized mortality ratio = observed deaths/expected deaths."
    ),
    digits = 1
)
```

### 2021-08-04 analysis/tables.Rmd
Revert "fix: working directory for targets"
This reverts commit cac14f630848b078c39aee7344862d418fcdcb52.
feat: add table 1 and table observed vs expected mortality

```{r histograms_setup, message = FALSE}
library("targets")
library("skimr")

withr::with_dir(rprojroot::find_rstudio_root_file(), {
    tar_load(raw_data)
    tar_load(imp_data)
    tar_load(zlog_data)
})
```

@@

>>>>

```{r histograms_setup, message = FALSE}
library("targets")
library("skimr")
library("zlog")
tar_load(raw_data)
tar_load(imp_data)
tar_load(zlog_data)
```

@@

# Raw Data (imputed) with Zlog Colors

```{r tables_raw_zlog, results = "asis", echo = FALSE, message = FALSE}
k <- kableExtra::kable_classic(
    kableExtra::kbl(
        raw_data,
        digits = 2,
        escape = FALSE
    ),
    "basic"
)
for (i in grep("_[CEQS]$", colnames(raw_data)))
    k <- kableExtra::column_spec(k, i, background = zcol(zlog_data[[i]]))
k |> kableExtra::scroll_box(width = "100%", height = "500px")
```

# Table One

```{r tbl1, results = "asis", echo = FALSE, message = FALSE}
tar_read(tbl1)
```
